Kelp morphometric properties and the link to hydrodynamic modelling

### 2018-10-03 chapter_1.Rmd
minor changes to chapter 1

---
title: "Kelp morphometric properties and the link to hydrodynamic modelling"
author: "Ross Coppin"
date: "26 February 2017"
output:
  pdf_document:
    highlight: zenburn
    number_sections: yes
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  word_document:
    toc: yes
bibliography: LaTeX/mycollection.bib
---

@@

Waves have a number of charateristics which are dipected in figure ??, and is an idealised representation of what a wave is. The *amplitude* is the vertical distance from its midline or *equilibrium surface* or *still water level* to its highest point known as the *crest*. The equalibrium surface is the level the ocean would be if there were no waves, for a wave to form a disturbance must occur below or above this line. The *trough* is the same distance of the amplitude, however the measurement is taken from the equilibrium surface to the lowest point. Wave height is the vertical distance from crest to trough, and is equal to twice the the amplitude. The *wavelength* is the horizontal distance from a crest/trough to the next crest/trough respectively. It is important to note that the energy progpgating from a disturbance will not reach the ocean floor in deep-water environments. The depth below a wave where the water, and anything in the water, feels no motion or disturbance is known as the *wave base*. The wave base is calculated by descending vertical from the equalibrium surface by a value equal to halve the wavelength.

As mentioned previously, the water particles in the ocean move in a circular orbit and hence return to their orginal position. This is because waves in the ocean represent moving energy, not moving water.

##Hydrodynamic modelling

>>>>

---
title: "Kelp morphometric properties and the link to hydrodynamic modelling"
author: "Ross Coppin"
date: "26 February 2017"
output:
  word_document:
    toc: yes
  html_document:
    df_print: paged
    toc: yes
  pdf_document:
    highlight: zenburn
    number_sections: yes
    toc: yes
bibliography: LaTeX/mycollection.bib
---

@@

Waves have a number of charateristics which are dipected in figure ??, and is an idealised representation of what a wave is. The *amplitude* is the vertical distance from its midline or *equilibrium surface* or *still water level* to its highest point known as the *crest*. The equalibrium surface is the level the ocean would be if there were no waves, for a wave to form a disturbance must occur below or above this line. The *trough* is the same distance of the amplitude, however the measurement is taken from the equilibrium surface to the lowest point. Wave height is the vertical distance from crest to trough, and is equal to twice the the amplitude. The *wavelength* is the horizontal distance from a crest/trough to the next crest/trough respectively. It is important to note that the energy progpgating from a disturbance will not reach the ocean floor in deep-water environments. The depth below a wave where the water, and anything in the water, feels no motion or disturbance is known as the *wave base*. The wave base is calculated by descending vertical from the equalibrium surface by a value equal to halve the wavelength.The water particles in the ocean move in a circular orbit and hence return to their orginal position. This is because waves in the ocean represent moving energy, not moving water.

##Types of waves

There are a variety of waves that form in the ocean and all differ in terms of period or wavelength (??image??). The longest wave that can form in the ocean are known as *trans-tidal waves*, and are generated by fluctuations in magnitism between the Earth's crust and atmosphere. The magnetic push and pull between the Moon and the Sun creates waves with a slightly shorter wavelength, known as tides. Their period and wavelength can also range from a few hours to more than a day, and from a few hundred to a thousand kilometers respectively. *Storm surge* tend to have a slightly shorter wavelength and period compared to tides. When low atmospheric pressure systems and high wind speeds in a storm it elevates the ocean surface, generating storm surge, which may cause flooding in coastal areas as it approaches the coastline. *Tsunami's* are on the lower end of the scale, and are generated by earhquakes or submarine 'land' slides. Their random nature makes them difficult to predict and increase amplitude as the approach the coast. This can make them considerbly large along coastal areas and often cause immense damage and loss of life.

# Hydrodynamic modelling
